---
title: "Class06 R functions"
author: "Philip Dai Le"
date: "10/17/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This is my work flow for class 06 in **BIMM 143**: "Intro to bioinformatics".


```{r}
#this is to demo a code chunk. This is a code chunk in R
plot(1:10)
```

##Practice reading files (again...)
read.table() functions differ in parameters:
read.delim; uses sep="\t"
read.csv; uses sep=","
read.csv2; uses sep=";"
read.table; uses sep=" "

```{r}
read.table("test1.txt", sep=",", header=TRUE)
read.table("test2.txt")
read.table("test3.txt", sep="\t", header = FALSE)
read.csv 
read.table("https://bioboot.github.io/bimm143_S19/class-material/test2.txt", sep="$",header= TRUE)
```

```{r}
add <- function(x, y=1) {
 # Sum the input x and y
 x + y
}
```

Use ctrl+alt+I to insert code chunk
```{r}
add(1)
add(5,5) 
#overides default y value
#add(5,"barry"), add(5,5,5) 
#can't use a third value for Z cause not initiated and can't use non-numerical value
```

"Our wee function is vectorized :-)" - Barry Grant
```{r}
add(x=1, y=4)
add(1, 4)
add(1)
add( c(1, 2, 3) )
add( c(1, 2, 3), 4 )
#add(1, 2, 2), error due to third value not initialized 
#add(x=1, y=“b”), error due to y="b"
```


What does this code do?
df$a <- (df$a - min(df$a)) / (max(df$a) - min(df$a))
df$b <- (df$b - min(df$a)) / (max(df$b) - min(df$b))
df$c <- (df$c - min(df$c)) / (max(df$c) - min(df$c))
df$d <- (df$d - min(df$d)) / (max(df$a) - min(df$d)) 

Silent killers in code, trying to use "a" for df$b and df$d

Here the the intent is far more clear:
df$a <- rescale (df$a)
Purpose of code is clearer, reduce mistakes from copy/paste, updating code is easier, and reduce duplication and facilitate re-use


Start with code snippet to simplify and reduce calculatino duplication
Ex:x <- (x - min(x)) / (max(x) - min(x))
```{r}
rescale <- function(x) {
 rng <-range(x)
 (x - rng[1]) / (rng[2] - rng[1])
}


```

```{r}
rescale(1:10)
?range
range (1:10)
range (1:20)
y<-range(c(5,1,2,5,20))
y[1]
y[2]
c("a","b")

```

Test some
```{r}
rescale(c(1,2,NA,3,10))
```

```{r}
x<-c(1,2,NA,3,10)
rng<-range(x, na.rm = TRUE)
rng
```

```{r}
rescale2<- function(x) {
  rng<-range(x, na.rm = TRUE)
  (x-rng[1])/(rng[2]-rng[1])
}
```

```{r}
rescale2 (c(1,2,NA,3,10))
```

```{r}
rescale3 <- function(x, na.rm=TRUE, plot=FALSE) {
 rng <-range(x, na.rm=na.rm)
 print("Hello")
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 return(answer)
 print("is it me you are looking for?")
 if(plot) {
 plot(answer, typ="b", lwd=4)
 }
 print("I can see it in ...")
 return(answer)
}

```

```{r}
rescale3(1:10, plot = FALSE)
```

```{r}
rescale3(1:10, plot=TRUE)
```

Additional Info regarding header sizes
#This is H1
Creates small header

## This is H2
Creates large header

### This is H3
Creates medium header