---
title: "Class15"
author: "Philip Dai Le"
date: "11/14/2019"
output: github_document
---
##Lecture 15: Pathway Analysis

**Section 1. Differential Expression Analysis**

Loading in data package and initializing external files
```{r}
library(DESeq2) #loading data files
metaFile <-"GSE37704_metadata.csv" #assigning data files to a variable for easy use
countFile <-"GSE37704_featurecounts.csv"
```

Importing both metadata and countdata
```{r}
colData = read.csv(metaFile, row.names=1)
head(colData)

countData = read.csv(countFile, row.names = 1)
head(countData)

#notice how the data output are in different formats, different number of columns. Need to adjust data for proper merging
```

Adjusting countData and colData for merging. Used "-1" to remove the odd first $length col.
```{r}
countData <-as.matrix(countData[,-1])
head(countData)
```

Need to check if the data files line up in countData and metaData.  Focus on the id values of the colData metadata file. Use of the "all()" function allows all entry checking in a vector to see if they're TRUE
```{r}
colnames(countData)
colnames(colData)

#How to check for all entries? Check colnames of countData and rownames of colData
all(colnames(countData) == rownames(colData))

#results all returning TRUE = good

```

```{r}
#test for the check
all(c(T,F,T))
```

countData needs to be filtered out for the zero entries in all the experiments.
```{r}
#rowSums(countData) >0, loads only data greater than 0. !=0 denotes the values that are not 0 thus TRUE or FALSE
countData = countData [rowSums(countData)>0,]
head(countData)
```

Setting up DESeq2 to run DESeq pipeline on our data
```{r}
library(DESeq2)
dds = DESeqDataSetFromMatrix(countData = countData, colData = colData, design=~condition)

dds = DESeq(dds)

dds

#res = results(dds)
res = results(dds, contrast=c("condition", "hoxa1_kd", "control_sirna"))
summary(res)
```

Plotting volcano plot
```{r}
plot (res$log2FoldChange, -log(res$padj))
```

Adjusting and coloring the volcano plot
```{r}
#Making a color vector for all genes
mycols <- rep("gray", nrow(res))

#Coloring red for hte genes with absolute foldchange greater than 2
mycols [abs(res$log2FoldChange)>2]<- "red"

#Coloring blue with adjusted p-value less than 0.01 and absolute fold change greater than 2

inds<- ("P-value"<0.01) & (abs(res$log2FoldChange) > 2)
mycols [inds] <- "blue"

plot (res$log2FoldChange, -log(res$padj), col = mycols, xlab = "Log2(FoldChange)", ylab="-Log(P-value)")
```

```{r eval= FALSE}
#BiocManager :: install ("AnnotationDbi")
#BiocManager :: install("org.his.eg.db")

```


```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```

```{r}
columns(org.Hs.eg.db)

res$symbol = mapIds(org.Hs.eg.db,
                    keys=row.names(countData), 
                    keytype="ENSEMBL",
                    column="SYMBOL",
                    multiVals="first")

res$entrez = mapIds(org.Hs.eg.db,
                    keys=row.names(countData),
                    keytype="ENSEMBL",
                    column="ENTREZID",
                    multiVals="first")

#res$name =   mapIds(org.Hs.eg.db, keys=row.names(countData),keytype="ENSEMBL",column="NAMEID",multiVals="first")

head(res, 10)

```

##Section 2. Pathway Analysis
Kegg pathways analysis using pathview. First, Installed bioconductor package: pawthview, gage, and gageData. The loaded in pathview.
```{r}
library (pathview)
```


Loading in data packages and outputting data
```{r}
library(gage)
library(gageData)

data(kegg.sets.hs)
data(sigmet.idx.hs)

# Focus on signaling and metabolic pathways only
kegg.sets.hs = kegg.sets.hs[sigmet.idx.hs]

# Examine the first 3 pathways
head(kegg.sets.hs, 3)
```


```{r}
foldchanges = res$log2FoldChange
names(foldchanges) = res$entrez
head(foldchanges)
```



Running **gage** pathway analysis.
```{r}
keggres = gage(foldchanges, gsets=kegg.sets.hs)
```

```{r}
attributes(keggres)
```

```{r}
head(keggres$less)
```

```{r}
pathview(gene.data=foldchanges, pathway.id="hsa04110")
```


Changing the input arguments to pathview()
```{r}
# A different PDF based output of the same data
pathview(gene.data=foldchanges, pathway.id="hsa04110", kegg.native=FALSE)
```


```{r}
## Focus on top 5 upregulated pathways here for demo purposes only
keggrespathways <- rownames(keggres$greater)[1:5]

# Extract the 8 character long IDs part of each string
keggresids = substr(keggrespathways, start=1, stop=8)
keggresids
```

```{r}
pathview(gene.data=foldchanges, pathway.id=keggresids, species="hsa")
```



##Section 3 Gene Ontology (G0)
```{r}
data(go.sets.hs)
data(go.subs.hs)

# Focus on Biological Process subset of GO
gobpsets = go.sets.hs[go.subs.hs$BP]

gobpres = gage(foldchanges, gsets=gobpsets, same.dir=TRUE)

lapply(gobpres, head)
```



