---
title: "Class12Git"
output: github_document
---
##Lecture 12 

##Loading bio3d libary and reading in 1hsg file then assigning it to hiv for convenience.
```{r}
library(bio3d)
file.name<- get.pdb("1hsg.pdb")
hiv<-read.pdb(file.name)

hiv
```

> **Q1**: What is the name of the two non protein resid values in this structure? What does resid
correspond to and how would you get a listing of all reside values in this structure? 

The two non-protein resd values are: HOH (127) and MK1 (1) 

The residue corresponds to the water being a stabilizing factor for the protein and MK1 would be the drug ligand in 1hsg.


##Trimming the 1hsg pdb file to create protein and ligand only objects
```{r}
prot<-trim.pdb(hiv, "protein")
lig<-trim.pdb(hiv, "ligand")

prot
lig
```

##Alternative manner for trimming
```{r}
protein<-atom.select(hiv, "protein", value = TRUE)
write.pdb(protein,file="1hsg_protein.pdb")
protein

ligand<-atom.select(hiv,"ligand",value=TRUE)
write.pdb(ligand,file="1hsg_ligand.pdb")
```


Reading in ADT file for hsg
```{r}
read.all("1hsg_protein.pdbqt")
```

##2.3 Inspecting your docking results 
```{r}
library(bio3d)
res <-read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res,"results.pdb")
```

